A Test for White Noise in the Entropy-Complexity Plane

### 2021-10-28 Reports/Journal - Confidence Regions/IT-ConfidenceRegions.tex
Reorganized
Update IT-ConfidenceRegions.tex

\author[1]{Eduarda T.\ C.\ Chagas}
\author[2,3]{Marcelo Queiroz}
\author[4]{Osvaldo A.\ Rosso}
\author[1]{Heitor S.\ Ramos}
\author[2]{Cristopher G.\ S. Freitas}
\author[2]{Leonardo V.\ Pereira}
\author[4]{Alejandro C.\ Frery}

@@

\begin{abstract}
The \citeauthor{PermutationEntropyBandtPompe} methodology has been used successfully in the analysis of time series, 
by computing Information Theory descriptors of the histogram of ordinal patterns.
Such descriptors lie in a 2D manifold: the Entropy-Complexity plane.
This article proposes the first approach that uses the entropy-complexity plane to test the white noise hypothesis.
Our test is based on true white noise sequences obtained from physical devices.
The proposed methodology provides consistent results, 
managing to assess sequences of true random samples as random (adequate size), 
rejecting correlated sequences (sound power), and
capturing the randomness of generators previously analyzed in the literature.
\keywords{Random number generators, Bandt–Pompe approach, Entropy-complexity plane, Time Series, Information Theory}
\end{abstract}

@@

>>>>

\author[1]{Eduarda T.\ C.\ Chagas}
\author[2,3]{Marcelo Queiroz-Oliveira}
\author[4]{Osvaldo A.\ Rosso}
\author[1]{Heitor S.\ Ramos}
\author[2]{Cristopher G.\ S. Freitas}
\author[5]{Alejandro C.\ Frery}

@@

		\begin{abstract}
			The \citeauthor{PermutationEntropyBandtPompe} methodology has been used successfully in the analysis of time series. It consists of computing Information Theory descriptors of the histogram of ordinal patterns.
			Such descriptors lie in a 2D manifold: the Entropy-Complexity plane.
			This article proposes the first proposal of a test in the entropy-complexity plane for the white noise hypothesis.
			Our test is based on true white noise sequences obtained from physical devices.
			The proposed methodology provides consistent results, 
			managing to assess sequences of true random samples as random (adequate size), 
			rejecting correlated sequences (sound power), and
			capturing the randomness of generators previously analyzed in the literature.
			\keywords{Random number generators, Bandt–Pompe approach, Entropy-complexity plane, Time Series, Information Theory}
		\end{abstract}

@@

\subsection{The Entropy-Complexity Plane in the Literature}\label{Sec:BPApplications}

Initial works on the characterization of white noises with permutation entropy arose from the need to discriminate between them and chaotic maps \citep{rosso2013characterization, xiong2020complexity}.
Notice that the logistic map has been used for pseudorandom number generation; cf. \citet{AnalyzingLogisticMapPseudorandomNumberGeneratorsforPeriodicityInducedbyFinitePrecisionFloatingPointRepresentation} and the references therein.
It was found that the measure of statistical complexity was able to efficiently quantify the performance of pseudorandom number generators, expanding the possibilities of using Information Theory descriptors with the Bandt-Pompe symbolization \citep{larrondo2013statistical, gonzalez2005statistical}.

Table~\ref{Tab:Literature} presents a summary of the main works in the literature that perform analysis of non-chaotic algorithmic generators, according to their features in the $H\times C$ plane.
We also provide the length $T$ and embedding dimension $D$ of the time series under scrutiny.
The following algorithmic generators were analyzed:
\begin{itemize}
	\item Mother RNG, available in Marsaglia's website \citep[MOT,][]{marsaglia1994yet};
	\item Multiple with carry RNG \citep[MWC,][]{marsaglia1994yet};
	\item Combo RNG \citep[COM,][]{marsaglia1994yet};
	\item Lehmer RNG \citep[LEH,][]{payne1969coding};
	\item Fractional Brownian motion (fBm) and fractional Gaussian noise (fGn); see \citet{bardet2003generators};
	\item Colored noise with power spectrum $f^{-k}$ with $k \geq 0$~\citep{larrondo2012matlab};
	\item Linear Congruential Generator \citep[LCG,][]{knuth1997sorting}.
\end{itemize}

\begin{table}[hbt]
	\caption{Results of the main works of white noise sequences analysis in the $H \times C$ plane.}
	\label{Tab:Literature}
	\centering
	\begin{tabular}{llcccc b{5.4em}}
		\toprule
		Reference & PRNG & $T$ & $D$ & $H$ & $C$ & Considered random? \\ 
		\midrule
		\citeauthor{larrondo2013statistical} (\citeyear{larrondo2013statistical}) &  MOT & NA & 6 & $\cong 0.9969$ & $\cong 0$ & no\\
		\cmidrule(lr){1-7}
		\citeauthor{gonzalez2005statistical} (\citeyear{gonzalez2005statistical})  &  MWC & 65536 & NA & $\cong 1$ & $0.3$ & yes \\
		&  MOT & 65536 & NA & $\cong 1$ & $0.3$ & yes \\
		&  COM & 65536 & NA & $\cong 1$ & $0.05$ & yes\\
		\cmidrule(lr){1-7}
		\citeauthor{RandomNumberGeneratorsCausality} (\citeyear{RandomNumberGeneratorsCausality}) &  LEH & \num[scientific-notation=true]{5 e6} & 5 & $\cong 1$ & $10^{-4}$ & yes \\
		&  MOT & \num[scientific-notation=true]{5 e6} & 5 & $\cong 1$ & $10^{-4}$ & yes\\
		&  MWC & \num[scientific-notation=true]{5 e6} & 5 & $\cong 1$ & $10^{-4}$ & yes\\
		\cmidrule(lr){1-7}
		\citeauthor{rosso2013characterization} (\citeyear{rosso2013characterization}) &  LCG & \num[scientific-notation=true]{1 e7} & 6 & $0.997871$ & $0.005101$ & no\\
		\cmidrule(lr){1-7}
		\citeauthor{xiong2020complexity} (\citeyear{xiong2020complexity}) &  fGn & \num[scientific-notation=true]{2 e17} & 6 & $\cong 1$ & $\cong 0$ & yes \\
		& $f^{-k}$ & \num[scientific-notation=true]{2 e17} & 6 & $\cong 1$ & $\cong 0$ & yes \\
		\bottomrule
	\end{tabular}
\end{table}

None of these works provide $p$-values or hypothesis tests for their analysis.
The focus of the articles is finding the set of descriptors that best discriminates chaos from noise.
The authors make assessments about the randomness of sequence on ad hoc visual inspection of the point's location in the $H\times C$ plane.
Our work fills such a gap for finite sequences of white noise and proposes a methodology that can be extended to any other situation.
